\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{bm}
\usepackage{enumitem}
\usepackage{hyperref}

\title{A Primer on the Implemented Korea Spatial GE Model}
% [REVIEW FLAG \S7.1] A submission needs a real author/title. Left as a primer for now.
\author{}
\date{\today}

% Region set. The macro was renamed away from a name that reads as the reals;
% the set of regions renders as \mathcal{N}. The unused \E macro was removed (review 7.1).
\newcommand{\Nreg}{\mathcal{N}}
\newcommand{\Sct}{\mathcal{J}}

\begin{document}
\maketitle

\section{Purpose}
This note documents the quantitative model implemented in the \texttt{korea\_growth}
package. It is a background primer rather than a full theory paper: the goal is to
state the objects, equilibrium conditions, and numerical solution method that are
actually used by the code.

The implementation solves a sequence of static spatial equilibria. The only dynamic
state is lagged population, so period $t$ equilibrium takes $L_{t-1}$ as given and
returns $(w_t,r_t,P_t,E_t,L_t,\bar\tau_t,\bar\pi_t)$.

\section{Environment}
Time is discrete, $t = 1,\dots,T$. Regions are indexed by $o,d \in \Nreg$ and sectors by
$j,k \in \Sct$. Let $N = |\Nreg|$ and $J = |\Sct|$.

The exogenous inputs in period $t$ are:
\begin{itemize}[leftmargin=1.5em]
\item regional amenities $\bar V_{d,t}$,
\item domestic trade costs $\tau_{odj,t} \ge 1$ and import costs $\tilde\tau_{oj,t} \ge 1$,
\item migration costs $\delta_{od,t} \ge 1$,
\item sectoral productivities $A_{oj,t}$,
\item land shares $\beta_{oj,t}$ and value-added shares $\gamma_{oj,t}$,
\item input-output shares $\gamma^{IO}_{ojk,t}$,
\item fixed costs $F_{oj,t}$, export fixed costs $\tilde F_{oj,t}$, and adoption fixed costs
      $\breve F_{oj,t}$,
\item subsidy rates $s_{oj,t}$,
\item foreign demand shifters $\tilde D_{j,t}$ and foreign prices $\tilde p_{j,t}$,
\item land or structures supply $H_{o,t}$,
\item the mass of potential firms $M_{j,t}$.
\end{itemize}

For the agriculture sector there is an alternative mechanized technology with
$(\breve\beta_{oj,t}, \breve\gamma_{oj,t}, \breve\gamma^{IO}_{ojk,t})$ and efficiency gain
$\xi > 1$.

The structural parameters are:
\[
\sigma > 1,\quad \theta > 0,\quad \kappa > 1,\quad \rho_j \in (0,1),\quad
\iota \in (-1,0),\quad \nu > 0,\quad \eta \in \mathbb{R},
\]
together with expenditure parameters $\alpha_j$ and $v_j$ satisfying
$\sum_j \alpha_j = 1$ and $\sum_j v_j = 0$.

\paragraph{Timing within a period (review \S7.5).} The within-period sequence is: (i)
lagged population $L_{t-1}$ determines agglomeration $f_{oj,t}=L_{o,t-1}^{\rho_j}$ and
congestion $g_{d,t}=L_{d,t-1}^{\iota}$; (ii) agents observe current-period
$(w_t,P_t,r_t,\bar\tau_t,\bar\pi_t)$ and migrate, so migration is \emph{myopic} on
current-period fundamentals; (iii) production, goods-market clearing, and consumption
occur in the destination, with the labor market clearing at the current-period population
$L_t$ (congestion is dated $t-1$, the active labor force is dated $t$). This myopia is a
modeling choice; forward-looking migration is a Phase-2 upgrade (review \S3.2).
% [REVIEW FLAG \S3.2] Myopic migration on unevenly spaced periods (1965,68,73,75,80,85)
% gives per-period elasticities no consistent annual interpretation. Annualization plus a
% forward-looking robustness check is a structural (non-theory-edit) item, flagged for the
% later phase, not fixed here.

\section{Households}
\subsection{Ownership, numeraire, and disposable income}
\paragraph{Numeraire (review \S1.5).} The economy is a small open economy that faces an
exogenous foreign price vector $\tilde p_{j,t}$. We fix the numeraire by normalizing the
foreign price of the heavy-manufacturing good, $\tilde p_{\mathrm{HeavyMnf},t}\equiv 1$;
all wages, rents, and price indices are expressed in these units. Because $L_{0}$ is
normalized to sum to one, all ``populations'' $L_{o,t}$ are population \emph{shares} and
wages $w_{o,t}$ are per-share; a bridge to won-denominated data therefore requires an
explicit scale factor (review \S7.10).
% [REVIEW FLAG \S1.5] Choice of numeraire (foreign HeavyMnf price = 1) is a modeling
% decision; an alternative is to fix aggregate nominal income W_t = 1. Confirm preference.

\paragraph{Ownership structure (review \S1.1, \S3.4).} We state ownership explicitly so
that income equals expenditure (\S\ref{sec:walras}):
\begin{itemize}[leftmargin=1.5em]
\item \emph{Labor} is supplied inelastically by residents; region-$d$ residents earn
      $w_{d,t}L_{d,t}$.
\item \emph{Land/structures} are owned locally: the land rent $r_{o,t}H_{o,t}$ generated
      in region $o$ accrues to the residents of $o$ and is rebated per capita. (Robustness
      to absentee/national ownership is a one-line change in the rebate rule.)
\item \emph{Firm profits} are pooled nationally and rebated in proportion to wage income,
      through the economy-wide rebate rate $\bar\pi_t$ defined in \S\ref{sec:gov}.
\end{itemize}
% [REVIEW FLAG \S1.1/\S3.4] Local land ownership and wage-proportional profit pooling are
% both distributionally loaded assumptions. The paper's incidence question may want
% local profit retention instead; flagged for a robustness pass.

\paragraph{Disposable income.} Per-capita disposable income in region $d$ is
\begin{equation}
  y^{pc}_{d,t}
  = (1 - \bar\tau_t + \bar\pi_t)\, w_{d,t}
  + \frac{r_{d,t} H_{d,t}}{L_{d,t}},
  \label{eq:ypc}
\end{equation}
where $\bar\tau_t$ is the proportional labor-income tax that finances subsidies and
infrastructure and $\bar\pi_t$ is the aggregate profit rebate in wage units. The land-rent
term restores the land income that the previous specification computed
(to clear the land market) but paid to no one; because rural land-value capitalization is
the canonical welfare metric for market-access gains (Donaldson--Hornbeck 2016), omitting
it discarded exactly the incidence the paper studies.

Given sectoral price indices $P_{djt}$, define the composite price index
\begin{equation}
  \mathcal{P}_{d,t} = \prod_{j \in \Sct} P_{djt}^{\alpha_j}.
\end{equation}
The implemented indirect utility is
\begin{equation}
  V_{d,t} =
  \begin{cases}
    \dfrac{1}{\eta} \left( \dfrac{y^{pc}_{d,t}}{\mathcal{P}_{d,t}} \right)^{\eta}
    - \sum_j v_j \log P_{djt}, & \eta \ne 0, \\[1em]
    \log\!\left(\dfrac{y^{pc}_{d,t}}{\mathcal{P}_{d,t}}\right)
    - \sum_j v_j \log P_{djt}, & \eta = 0.
  \end{cases}
\end{equation}
This is the price-independent generalized-linear (PIGL) indirect utility of Boppart
(2014), which should be cited and whose regularity conditions the model inherits.

\paragraph{Regularity and sign conventions (review \S7.3).} For the expenditure shares
$\psi_{dj,t}$ of \S3.3 to lie in $[0,1]$ over the equilibrium income range, and for
Engel's law (rising income shifting demand away from agriculture) to hold, the parameters
must satisfy Boppart's conditions rather than merely $\eta\in\mathbb{R}$: we require
$\eta>0$ and $v_{\mathrm{Agri}}>0$ (so the agricultural share $\alpha_{\mathrm{Agri}} +
v_{\mathrm{Agri}}(y^{pc}/\mathcal{P})^{-\eta}$ declines in real income), together with
income bounds guaranteeing $0\le \alpha_j + v_j (y^{pc}/\mathcal{P})^{-\eta}\le 1$ for
every $j$. These restrictions should be checked at the solved equilibrium, not merely
imposed on primitives.
% [REVIEW FLAG \S7.3] The current code permits eta of either sign and does not verify the
% share-bound conditions at the solution. Enforcing eta>0, v_Agri>0, and the share bounds
% is a checks.py addition, deferred to the code pass.

\subsection{Migration}\label{sec:migration}
\paragraph{Shock structure (review \S2.3, \S7.2).} The previous draft wrote migration
shares as $\mu_{od}\propto(\bar V_d L_{d,t-1}^{\iota} V_d/\delta_{od})^{\nu}$ and called
this ``multinomial logit in levels.'' That is a mislabel --- it is a power (Fr\'echet-type)
share --- and, more seriously, it is ill-posed: with the PIGL indirect utility $V_d$ of
\S3.1 negative (which happens for $\eta<0$, or for $y^{pc}<\mathcal{P}$ in the log case),
$V_d^{\nu}$ is not real and dividing by $\delta_{od}\ge 1$ \emph{raises} utility. We
re-derive the block from an explicit shock structure so that it is well-defined for all
parameter values and delivers a welfare formula (\S\ref{sec:welfare}).

An agent $i$ resident in $o$ draws idiosyncratic taste shocks $\{\varepsilon_{id}\}_{d\in\Nreg}$,
i.i.d.\ Gumbel with scale $1/\nu$, and chooses the destination maximizing the additively
separable value
\begin{equation}
  W_{od,t} + \tfrac{1}{\nu}\,\varepsilon_{id},
  \qquad
  W_{od,t}
  = \underbrace{\log \bar V_{d,t}}_{\text{amenity}}
  + \underbrace{\iota \log L_{d,t-1}}_{\text{congestion}}
  + \underbrace{\log \mathcal{Y}_{d,t}}_{\text{real income}}
  - \underbrace{\log \delta_{od,t}}_{\text{migration cost}} ,
  \label{eq:Wod}
\end{equation}
where $\mathcal{Y}_{d,t}>0$ is a money-metric real-income index for destination $d$ (the
level of income, at a fixed reference price vector, that delivers the household's
attained utility). Since PIGL utility is defined only up to affine transformations,
migration elasticities must be built on this expenditure-function-based measure rather
than on the cardinal level $V_d$; that is what removes the cardinalization dependence the
referee flags. Operationally we use the log real-income index
\begin{equation}
  \log \mathcal{Y}_{d,t}
  = \log\!\left(\frac{y^{pc}_{d,t}}{\mathcal{P}_{d,t}}\right)
  - \sum_{j} v_j \log P_{djt},
  \label{eq:realincome}
\end{equation}
which is the money-metric welfare of the PIGL system in the $\eta\to 0$ (log) case and a
monotone transform of $V_{d,t}$ in general; it is finite and well-defined for all
$y^{pc}_{d,t}>0$.
% [REVIEW FLAG \S2.3] Eq. (realincome) is the exact money metric only in the log-preference
% case. For general eta the exact PIGL expenditure function should replace it (appendix
% derivation, inheriting Boppart 2014 regularity conditions). Flagged as a to-derive item.

\paragraph{Choice probabilities and population dynamics.} With Gumbel shocks the
destination probabilities are the standard logit
\begin{equation}
  \mu_{od,t}
  = \frac{\exp(\nu\, W_{od,t})}{\sum_{d'\in\Nreg}\exp(\nu\, W_{od',t})}
  = \frac{\bigl(\bar V_{d,t}\, L_{d,t-1}^{\iota}\, \mathcal{Y}_{d,t}/\delta_{od,t}\bigr)^{\nu}}
         {\sum_{d'\in\Nreg}\bigl(\bar V_{d',t}\, L_{d',t-1}^{\iota}\, \mathcal{Y}_{d',t}/\delta_{od',t}\bigr)^{\nu}},
  \label{eq:mu}
\end{equation}
which is manifestly in $[0,1]$ and well-defined because $\mathcal{Y}_{d,t}>0$. The second
equality shows the earlier power form is recovered \emph{once the positive real-income
index $\mathcal{Y}_{d,t}$ replaces the sign-ambiguous $V_{d,t}$}. Population evolves as
\begin{equation}
  L_{d,t} = \sum_{o \in \Nreg} \mu_{od,t}\, L_{o,t-1}.
\end{equation}

\subsection{Non-homothetic final demand}
Expenditure shares in destination $d$ are
\begin{equation}
  \psi_{dj,t}
  = \alpha_j + v_j \left( \frac{y^{pc}_{d,t}}{\mathcal{P}_{d,t}} \right)^{-\eta}.
\end{equation}
Final consumption expenditure is
\begin{equation}
  E^{F}_{dj,t} = \psi_{dj,t} y^{pc}_{d,t} L_{d,t}.
\end{equation}

\section{Production and firm heterogeneity}
\subsection{Unit costs and agglomeration}
Agglomeration is sector specific:
\begin{equation}
  f_{oj,t} = L_{o,t-1}^{\rho_j}.
\end{equation}

For the baseline technology, unit cost in region $o$ and sector $j$ is
\begin{equation}
  c_{oj,t}
  =
  \left[
    \frac{1}{\gamma_{oj,t}}
    \left( \frac{r_{o,t}}{\beta_{oj,t}} \right)^{\beta_{oj,t}}
    \left( \frac{w_{o,t}}{1-\beta_{oj,t}} \right)^{1-\beta_{oj,t}}
  \right]^{\gamma_{oj,t}}
  \prod_{k \in \Sct}
  \left( \frac{P_{ok,t}}{\gamma^{IO}_{ojk,t}} \right)^{\gamma^{IO}_{ojk,t}}.
\end{equation}
The mechanized agricultural technology has an analogous unit cost
$\breve c_{oj,t}$ built from $(\breve\beta,\breve\gamma,\breve\gamma^{IO})$.

\subsection{Productivity draws}
Potential entrants draw idiosyncratic productivity $\phi \in [1,\kappa]$ from a bounded
Pareto distribution with cdf
\begin{equation}
  G(\phi) = \frac{1 - \phi^{-\theta}}{1 - \kappa^{-\theta}},
  \qquad \phi \in [1,\kappa],
\end{equation}
and density
\begin{equation}
  g(\phi) = \frac{\theta \phi^{-\theta-1}}{1 - \kappa^{-\theta}}.
\end{equation}
The main integral used by the code is
\begin{equation}
  I(a,b) = \int_a^b \phi^{\sigma-1} g(\phi)\, d\phi.
\end{equation}

\subsection{Demand shifter}
Given destination expenditures and prices, origin $o$ faces sector-$j$ demand shifter
\begin{equation}
  S_{oj,t}
  = \sum_{d \in \Nreg} \tau_{odj,t}^{1-\sigma} P_{djt}^{\sigma-1} E_{dj,t}.
\end{equation}
Here $E_{dj,t}$ is \emph{total} sector-$j$ expenditure in destination $d$ (final demand
plus intermediate demand from roundabout production), not final demand alone; this is the
correct object under input--output linkages (review \S7.7).

\subsection{Entry, adoption, and export cutoffs}\label{sec:cutoffs}
Each cutoff is the productivity at which the marginal firm's operating profit just covers
the relevant fixed cost; the $\sigma^{1/(\sigma-1)}$ factor comes from substituting the
CES markup pricing into that zero-cutoff-profit condition (this and every cutoff formula
below should be derived, not asserted --- review \S6, \S7). The zero-profit entry cutoff is
\begin{equation}
  \bar\phi_{oj,t}
  =
  \max\left\{
    1,\,
    \frac{\sigma}{\sigma-1}
    \frac{1-s_{oj,t}}{A_{oj,t} f_{oj,t}}
    \sigma^{\frac{1}{\sigma-1}}
    c_{oj,t}
    \left( \frac{F_{oj,t}}{S_{oj,t}} \right)^{\frac{1}{\sigma-1}}
  \right\}.
\end{equation}
The $\max\{1,\cdot\}$ truncation binds when the unconstrained cutoff falls below the lower
support point: economically, \emph{all} potential entrants are then active and the
zero-cutoff-profit condition fails at an interior $\phi$ (the marginal drawable firm earns
strictly positive profit). The map has a kink there; \S\ref{sec:existence} discusses the
consequences (review \S7.8, \S2.4).

For agriculture, the adoption cutoff uses the mechanization benefit term
\begin{equation}
  B_{oj,t}
  =
  \left[
    \left( \xi \frac{c_{oj,t}}{\breve c_{oj,t}} \right)^{\sigma-1} - 1
  \right]^{\frac{1}{1-\sigma}},
\end{equation}
whenever the expression in brackets is positive. Then
\begin{equation}
  \breve\phi_{oj,t}
  =
  \max\left\{
    1,\,
    \frac{\sigma}{\sigma-1}
    \frac{1-s_{oj,t}}{A_{oj,t} f_{oj,t}}
    \sigma^{\frac{1}{\sigma-1}}
    c_{oj,t} B_{oj,t}
    \left( \frac{\breve F_{oj,t}}{S_{oj,t}} \right)^{\frac{1}{\sigma-1}}
  \right\}.
\end{equation}
If the benefit term is non-positive, the code sets $\breve\phi_{oj,t} = \infty$ and
mechanized adoption never occurs.

The export cutoff is
\begin{equation}
  \tilde\phi_{oj,t}
  =
  \max\left\{
    1,\,
    \frac{\sigma}{\sigma-1}
    \frac{1-s_{oj,t}}{A_{oj,t} f_{oj,t}}
    \sigma^{\frac{1}{\sigma-1}}
    c_{oj,t}
    \left(
      \frac{\tilde F_{oj,t}}
      {\tilde\tau_{oj,t}^{1-\sigma} \tilde D_{j,t}}
    \right)^{\frac{1}{\sigma-1}}
    \times \chi_{oj,t}
  \right\},
\end{equation}
where $\chi_{oj,t} = (\breve c_{oj,t}/c_{oj,t}) / \xi$ for agriculture and
$\chi_{oj,t} = 1$ otherwise.

\paragraph{Cutoff ordering (review \S2.2).} For agriculture the export term carries
$\chi=(\breve c/c)/\xi$, i.e.\ the export cutoff is written as though every exporter
already operates the mechanized technology, yet nothing in the current implementation
enforces $\tilde\phi_{oj,t}\ge \breve\phi_{oj,t}$. With cheap exporting or expensive
adoption the model can therefore manufacture ``mechanized exporters'' below the adoption
cutoff --- firms that never paid $\breve F$ and are simultaneously counted as traditional
producers domestically. Two coherent resolutions:
\begin{enumerate}[leftmargin=1.5em]
\item \emph{Restrict and verify.} State primitive conditions guaranteeing the ordering
      $\bar\phi_{oj,t}\le\breve\phi_{oj,t}\le\tilde\phi_{oj,t}$ and enforce it as a check.
\item \emph{Solve the joint discrete choice.} Model the adoption--export decision jointly
      (exporting raises the return to adopting, a genuine complementarity) rather than
      pricing the two margins as if independent.
\end{enumerate}
We adopt (1) as the near-term fix and record the ordering as an equilibrium restriction in
\S\ref{sec:restrictions}; (2) is a Phase-2 modeling upgrade.
% [REVIEW FLAG \S2.2] Resolution (1) presumes the ordering holds at the calibrated
% parameters. This must be checked in checks.py (code pass) and may fail for some
% policy configurations, in which case (2) is required. Flagged.

\subsection{Productivity aggregates}
An adopter with idiosyncratic draw $\phi$ produces with effective productivity
$\xi\phi$ \emph{in every market}: that is precisely the premise of the adoption benefit
term $B_{oj,t}$ (\S\ref{sec:cutoffs}), which selects the adoption cutoff on the assumption of the $\xi$
gain. The mechanized productivity aggregates must therefore carry the same $\xi$
domestically as they do in the export block. Define
\begin{align}
  Z_{oj,t} &= A_{oj,t} f_{oj,t}\, I(\bar\phi_{oj,t}, \bar b_{oj,t})^{\frac{1}{\sigma-1}}, \\
  \breve Z_{oj,t} &= A_{oj,t} f_{oj,t}\, \xi\, I(\breve\phi_{oj,t}, \kappa)^{\frac{1}{\sigma-1}},
  \label{eq:Zbreve}\\
  \tilde Z_{oj,t} &= A_{oj,t} f_{oj,t}\, \chi^x_j\, I(\tilde\phi_{oj,t}, \kappa)^{\frac{1}{\sigma-1}},
\end{align}
where $\bar b_{oj,t} = \min\{\breve\phi_{oj,t}, \kappa\}$ in agriculture and
$\bar b_{oj,t} = \kappa$ otherwise, and $\chi^x_j = \xi$ for agriculture and $1$ otherwise.

\paragraph{Correction (review \S2.1).} The $\xi$ factor in \eqref{eq:Zbreve} was
\emph{absent} in the previous draft and in \texttt{trade.py}: adopters were selected on a
cutoff that assumes the $\xi$ gain, paid the full adoption fixed cost $\breve F$, and were
then aggregated into the domestic price index and domestic revenues \emph{without} the
gain, understating both the price-index benefit of mechanization and mechanized domestic
revenue by a factor $\xi^{\sigma-1}$. Because agricultural mechanization is the paper's
headline rural channel, this directly attenuates the mechanism the model exists to
quantify; the fix is to include $\xi$ in $\breve Z_{oj,t}$ exactly as in $\tilde Z_{oj,t}$.
% [REVIEW FLAG \S2.1] This is a THEORY correction; the corresponding one-line code fix
% (multiply Zbreve by xi in trade.py:203) is deferred to the code pass and WILL change all
% numerical outputs and may break directional tests. See discrepancies section.

\section{Sectoral price indices and revenues}
\subsection{Price indices}
Let
\begin{equation}
  \Omega_{oj,t}
  =
  M_{j,t}
  \left( \frac{\sigma}{\sigma-1}(1-s_{oj,t}) \right)^{1-\sigma}.
\end{equation}
The baseline and mechanized domestic supply terms are
\begin{align}
  B^{trad}_{oj,t} &= \Omega_{oj,t} \left( \frac{Z_{oj,t}}{c_{oj,t}} \right)^{\sigma-1}, \\
  B^{new}_{oj,t} &=
  \begin{cases}
    \Omega_{oj,t} \left( \dfrac{\breve Z_{oj,t}}{\breve c_{oj,t}} \right)^{\sigma-1},
    & \text{agriculture}, \\
    0, & \text{otherwise}.
  \end{cases}
\end{align}
Then the destination-$d$ price index in sector $j$ satisfies
\begin{equation}
  P_{djt}^{1-\sigma}
  =
  \sum_{o \in \Nreg} \left( B^{trad}_{oj,t} + B^{new}_{oj,t} \right) \tau_{odj,t}^{1-\sigma}
  + (\tilde\tau_{dj,t} \tilde p_{j,t})^{1-\sigma}.
\end{equation}

\subsection{Domestic and foreign revenues}
Domestic revenues are
\begin{align}
  R^{trad}_{oj,t} &= B^{trad}_{oj,t} S_{oj,t}, \\
  R^{new}_{oj,t} &= B^{new}_{oj,t} S_{oj,t}, \\
  R^{dom}_{oj,t} &= R^{trad}_{oj,t} + R^{new}_{oj,t}.
\end{align}

Export revenue is
\begin{equation}
  R^{x}_{oj,t}
  =
  M_{j,t}
  \left(
    \frac{\sigma}{\sigma-1}
    (1-s_{oj,t}) \tilde\tau_{oj,t}
  \right)^{1-\sigma}
  \left( \frac{\tilde Z_{oj,t}}{c^x_{oj,t}} \right)^{\sigma-1}
  \tilde D_{j,t},
\end{equation}
where $c^x_{oj,t} = \breve c_{oj,t}$ for agriculture and $c^x_{oj,t} = c_{oj,t}$ otherwise.
Gross output is
\begin{equation}
  Y_{oj,t} = R^{dom}_{oj,t} + R^{x}_{oj,t}.
\end{equation}

The mass of active firms, exporters, and adopters is
\begin{align}
  N^{firm}_{oj,t} &= M_{j,t} [1 - G(\bar\phi_{oj,t})], \\
  N^{x}_{oj,t} &= M_{j,t} [1 - G(\tilde\phi_{oj,t})], \\
  N^{adopt}_{oj,t} &=
  \begin{cases}
    M_{j,t} [1 - G(\breve\phi_{oj,t})], & \text{agriculture}, \\
    0, & \text{otherwise}.
  \end{cases}
\end{align}

\section{Goods markets and factor markets}\label{sec:factor}
\subsection{Subsidy accounting: private cost versus factor payments (review \S1.3)}
The subsidy $s_{oj,t}$ enters prices as $p = \frac{\sigma}{\sigma-1}(1-s_{oj,t})\,c/\phi$,
so it lowers the firm's \emph{effective} marginal cost to $(1-s_{oj,t})c/\phi$ while the
\emph{true} resource cost per unit remains $c/\phi$. Three distinct objects must be kept
apart; conflating them was the source of the accounting leak:
\begin{align}
  \text{true factor cost (resource payment):}\quad
    TFC_{oj,t} &= \frac{\sigma-1}{\sigma\,(1-s_{oj,t})}\,Y_{oj,t}, \label{eq:TFC}\\
  \text{firm's private outlay:}\quad
    (1-s_{oj,t})\,TFC_{oj,t} &= \frac{\sigma-1}{\sigma}\,Y_{oj,t}, \\
  \text{government subsidy outlay:}\quad
    s_{oj,t}\,TFC_{oj,t} &= \frac{\sigma-1}{\sigma}\,\frac{s_{oj,t}}{1-s_{oj,t}}\,Y_{oj,t}.
\end{align}
Operating profit is $Y_{oj,t}/\sigma$ in all cases (buyers pay $Y$, the firm collects the
subsidy and pays true factor cost), so the profit block of \S\ref{sec:gov} is unaffected. But
\emph{factor income and intermediate demand are the full $TFC$}, not the firm's private
$\frac{\sigma-1}{\sigma}Y$: the previous draft (and \texttt{equilibrium.py}) used
$\frac{\sigma-1}{\sigma}Y$ for both, understating factor demand and intermediate demand in
subsidized cells by the factor $1/(1-s_{oj,t})$ (33\% at the calibrated $s=0.25$ in
Changwon) and dropping the taxed subsidy revenue before it reached factor owners. We
therefore write the resource-payment split by technology as
\begin{align}
  TFC^{trad}_{oj,t} &= \frac{\sigma-1}{\sigma\,(1-s_{oj,t})}\,R^{trad}_{oj,t}, &
  TFC^{new}_{oj,t}  &= \frac{\sigma-1}{\sigma\,(1-s_{oj,t})}\,\left(R^{new}_{oj,t}+R^{x}_{oj,t}\right),
\end{align}
with $TFC_{oj,t}=TFC^{trad}_{oj,t}+TFC^{new}_{oj,t}$.
% [REVIEW FLAG \S1.3] Grossing up by 1/(1-s) changes every factor-market and government
% residual in the code (equilibrium.py:135,206). This is a THEORY correction; the code
% edit and a Walras unit test are deferred to the code pass.

\subsection{Expenditure system}
Local intermediate demand is the material-input share of the \emph{true factor cost}
(the seller of each material receives its full market value; the buyer's subsidy is a
transfer that does not reduce the seller's receipts):
\begin{equation}
  E^{M}_{oj,t} = \sum_{k \in \Sct} TFC_{ok,t}\, \gamma^{IO}_{okj,t},
\end{equation}
except in agriculture, where the baseline and mechanized input shares are applied to the
corresponding technology blocks,
\begin{equation}
  E^{M}_{oj,t}\big|_{k=a\ \text{term}}
  = TFC^{trad}_{oa,t}\, \gamma^{IO}_{oaj,t}
  + TFC^{new}_{oa,t}\, \breve\gamma^{IO}_{oaj,t}.
\end{equation}
Total expenditure is
\begin{equation}
  E_{oj,t} = E^{F}_{oj,t} + E^{M}_{oj,t} + E^{G}_{oj,t},
\end{equation}
where $E^{G}_{oj,t}$ is government infrastructure demand for good $j$ (\S\ref{sec:gov}): unlike the
previous draft, infrastructure spending \emph{buys output} rather than vanishing.

\subsection{Factor prices}
Labor and land income are the labor and land shares of the \emph{true factor cost}
$TFC$ (not the firm's private outlay), plus --- for labor --- the fixed costs, which are
paid in local labor rather than destroyed (review \S1.2). Let the regional fixed-cost labor
bill be
\begin{equation}
  \Phi^{F}_{o,t}
  = \sum_{j\in\Sct}\left(
      N^{firm}_{oj,t} F_{oj,t}
      + N^{x}_{oj,t} \tilde F_{oj,t}
      + N^{adopt}_{oj,t} \breve F_{oj,t}
    \right),
\end{equation}
which measures the resources absorbed by entry, exporting, and adoption. For
non-agriculture the value-added split is
\begin{align}
  w_{o,t} L_{o,t} &= \sum_j (1-\beta_{oj,t}) \gamma_{oj,t}\, TFC_{oj,t} \;+\; \Phi^{F}_{o,t}, \\
  r_{o,t} H_{o,t} &= \sum_j \beta_{oj,t} \gamma_{oj,t}\, TFC_{oj,t}.
\end{align}
For agriculture the baseline and mechanized blocks are split:
\begin{align}
  w_{o,t} L_{o,t}
  &=
  (1-\beta_{oa,t}) \gamma_{oa,t}\, TFC^{trad}_{oa,t}
  + (1-\breve\beta_{oa,t}) \breve\gamma_{oa,t}\, TFC^{new}_{oa,t} \notag\\
  &\quad
  + \sum_{j \ne a} (1-\beta_{oj,t}) \gamma_{oj,t}\, TFC_{oj,t}
  \;+\; \Phi^{F}_{o,t}, \\
  r_{o,t} H_{o,t}
  &=
  \beta_{oa,t} \gamma_{oa,t}\, TFC^{trad}_{oa,t}
  + \breve\beta_{oa,t} \breve\gamma_{oa,t}\, TFC^{new}_{oa,t}
  + \sum_{j \ne a} \beta_{oj,t} \gamma_{oj,t}\, TFC_{oj,t}.
\end{align}
% [REVIEW FLAG \S1.2] Fixed costs are here modeled as demand for local labor (denominated
% in the numeraire), so they are wage income and are simultaneously netted from profits in
% \S6 --- a transfer, not a leak. An alternative is to make F a labor-quantity requirement
% priced at w_o (so the fixed-cost bill scales with local wages). The former keeps F
% comparable to its current numeraire units; confirm the intended units (review \S7.9).

\section{Government budget and profit rebate}\label{sec:gov}
Aggregate wage income is
\begin{equation}
  W_t = \sum_{o \in \Nreg} w_{o,t} L_{o,t}.
\end{equation}
Aggregate profits are
\begin{equation}
  \Pi_t
  =
  \sum_{o \in \Nreg} \sum_{j \in \Sct}
  \left[
    \frac{1}{\sigma} Y_{oj,t}
    - N^{firm}_{oj,t} F_{oj,t}
    - N^{x}_{oj,t} \tilde F_{oj,t}
    - N^{adopt}_{oj,t} \breve F_{oj,t}
  \right].
\end{equation}
Here the fixed costs are subtracted from operating profit \emph{and} appear as labor
income in \S\ref{sec:factor}, so they are a transfer rather than a leak. The profit rebate is
\begin{equation}
  \bar\pi_t = \frac{\Pi_t}{W_t}.
\end{equation}
Aggregate profit $\Pi_t$ can be \emph{negative} in a period where fixed costs exceed
operating profit; then $\bar\pi_t<0$ and disposable income \eqref{eq:ypc} is reduced. The
solver must accommodate this (e.g.\ by updating $1+\bar\pi_t$ in levels) rather than
flooring $\bar\pi_t$ at a small positive number and reporting a floored point as converged
(review \S2.4).
% [REVIEW FLAG \S2.4] solver.py/equilibrium.py currently floor pibar (and taubar) at
% eps=1e-14 in log space, which silently projects negative-profit states to ~0. Fixing the
% update to levels is a code-pass item.

Government spending is financed by the labor-income tax $\bar\tau_t$. Using the correct
subsidy outlay \eqref{eq:TFC} and letting infrastructure \emph{purchase goods},
\begin{equation}
  \bar\tau_t
  =
  \frac{
    \displaystyle\sum_{o,j} s_{oj,t}\, TFC_{oj,t}
    + G^{infra}_t
  }{W_t}
  =
  \frac{
    \displaystyle\sum_{o,j} \frac{\sigma-1}{\sigma}\,\frac{s_{oj,t}}{1-s_{oj,t}}\, Y_{oj,t}
    + G^{infra}_t
  }{W_t},
\end{equation}
where the subsidy term now carries the $1/(1-s_{oj,t})$ factor missing from the previous
draft (review \S1.3). Infrastructure spending $G^{infra}_t$ is allocated across sectors as
final-type demand $E^{G}_{oj,t}$ (e.g.\ construction/heavy-manufacturing and services), so
it enters goods-market clearing instead of being a pure tax:
\begin{equation}
  \sum_{o\in\Nreg}\sum_{j\in\Sct} E^{G}_{oj,t} = G^{infra}_t .
\end{equation}
The ``roads'' counterfactual then combines a trade-cost reduction with the fiscal cost of
building it, rather than a free trade-cost reduction plus an unrelated lump-sum tax.
% [REVIEW FLAG \S1.4] The sectoral allocation of G^infra (which goods roads/buildings buy)
% is a modeling choice; a natural default is heavy-manufacturing + services in fixed
% shares, or a dedicated construction sector. Confirm the intended mapping. The current
% code has no E^G at all (infra is a pure tax), so this needs a code change.

\section{Trade balance, adding-up, and welfare}\label{sec:walras}
\subsection{Trade balance (review \S1.5)}
Region-$o$ exports earn $\sum_j R^{x}_{oj,t}$; the economy's imports are the CES
expenditure allocated to the foreign variety in each sector and destination,
\begin{equation}
  \mathrm{IM}_t
  = \sum_{d\in\Nreg}\sum_{j\in\Sct}
    \frac{(\tilde\tau_{dj,t}\,\tilde p_{j,t})^{1-\sigma}}{P_{djt}^{1-\sigma}}\, E_{dj,t},
  \qquad
  \mathrm{EX}_t = \sum_{o\in\Nreg}\sum_{j\in\Sct} R^{x}_{oj,t}.
\end{equation}
The aggregate budget must state whose income absorbs any gap. We treat this as a small
open economy with an exogenous net foreign transfer $\mathrm{NX}^{\star}_t$ (default zero,
i.e.\ balanced trade):
\begin{equation}
  \mathrm{EX}_t - \mathrm{IM}_t = \mathrm{NX}^{\star}_t .
  \label{eq:tradebalance}
\end{equation}
A nonzero $\mathrm{NX}^{\star}_t$ (a trade deficit financed from abroad) must be added to
aggregate household income for the accounts to close.
% [REVIEW FLAG \S1.5] The current code imposes no trade-balance condition and no numeraire
% statement; exports are driven by the exogenous shifter Dtilde and imports by CES demand,
% with nothing linking them. Enforcing (tradebalance) --- or explicitly carrying a deficit
% into income --- is a code-pass decision. Balanced trade is the cleanest default; confirm.

\subsection{Aggregate resource constraint and Walras's law}
With the corrections above --- land rents paid to residents \eqref{eq:ypc}, fixed costs
paid in labor, subsidies scaled by $1/(1-s)$, and infrastructure buying goods --- aggregate
income equals aggregate expenditure. Writing aggregate disposable income as
\begin{equation}
  Y^{D}_t
  = \underbrace{W_t}_{\text{wages}}
  + \underbrace{\textstyle\sum_o r_{o,t}H_{o,t}}_{\text{land rents}}
  + \underbrace{\Pi_t}_{\text{profits}}
  - \underbrace{\bar\tau_t W_t}_{\text{taxes}}
  + \underbrace{\mathrm{NX}^{\star}_t}_{\text{foreign transfer}},
\end{equation}
household budget balance gives $\sum_{o,j} E^{F}_{oj,t} = Y^{D}_t$, and goods-market
clearing (each variety's output equals domestic final $+$ intermediate $+$ government $+$
export demand for it) then closes the system. Equivalently, the model must satisfy the
single aggregate resource constraint
\begin{equation}
  \sum_{o,j} Y_{oj,t}
  = \sum_{o,j}\left( E^{F}_{oj,t} + E^{M}_{oj,t} + E^{G}_{oj,t}\right)
  + \mathrm{EX}_t - \mathrm{IM}_t
  \quad\text{(net of iceberg losses)} .
  \label{eq:walras}
\end{equation}
This adding-up should be \emph{proven} for the stated ownership structure and
\emph{unit-tested} at the solved equilibrium to numerical tolerance; every referee who
checks the accounting will look for it (review \S1, closing note).
% [REVIEW FLAG \S1] Eq. (walras) is the target identity, stated at the level of a primer.
% A rigorous proposition (income = expenditure, with the exact iceberg bookkeeping) belongs
% in the paper (Phase 6). The Walras unit test is a code-pass deliverable.

\subsection{Welfare (review \S3.9)}\label{sec:welfare}
The Gumbel shock structure of \S\ref{sec:migration} delivers a welfare metric, which the
previous draft lacked. Before observing idiosyncratic tastes, the expected utility of an
agent resident in origin $o$ is the log-sum (inclusive value)
\begin{equation}
  \mathbb{E}\,U_{o,t}
  = \frac{1}{\nu}\log\!\left(\sum_{d\in\Nreg}\exp(\nu\, W_{od,t})\right)
  + \frac{\gamma_{\mathrm{E}}}{\nu},
  \label{eq:welfare}
\end{equation}
where $W_{od,t}$ is the deterministic value in \eqref{eq:Wod} and $\gamma_{\mathrm{E}}$ is
the Euler--Mascheroni constant. Because $W_{od,t}$ is built on the money-metric real-income
index $\mathcal{Y}_{d,t}$, \eqref{eq:welfare} is a valid, cardinalization-robust welfare
measure. Aggregate welfare is the population-weighted sum $\sum_o L_{o,t-1}\,\mathbb{E}U_{o,t}$;
distributional incidence --- welfare by origin region, and by landowner versus worker and
by income level (through the PIGL shares) --- is the object the paper is built to deliver,
and is where this framework improves on aggregate HCI evaluations.
% [REVIEW FLAG \S3.9] The exact money-metric for general eta (\S2.3 flag) feeds directly into
% this welfare number; the two must use the same expenditure function. Deriving both is a
% Phase-1 theory deliverable.

\section{Static equilibrium}
Given $L_{t-1}$ and exogenous period-$t$ fundamentals, a static equilibrium is a tuple
\[
  \left\{
    w_{o,t}, r_{o,t}, P_{oj,t}, E_{oj,t}, L_{o,t}, \bar\tau_t, \bar\pi_t
  \right\}_{o \in \Nreg,\, j \in \Sct}
\]
such that (review \S7.6, a self-contained definition):
\begin{enumerate}[leftmargin=1.5em]
\item \textbf{Migration and population.} $\mu_{od,t}$ solves \eqref{eq:mu} given
      $(w,P,\bar\tau,\bar\pi,r,H)$ through disposable income \eqref{eq:ypc} and the
      real-income index \eqref{eq:realincome}, and $L_{d,t}=\sum_o\mu_{od,t}L_{o,t-1}$.
\item \textbf{Final demand.} $E^{F}_{dj,t}=\psi_{dj,t}\,y^{pc}_{d,t}L_{d,t}$ with the
      PIGL shares $\psi_{dj,t}$.
\item \textbf{Firm behavior.} Cutoffs $\bar\phi,\breve\phi,\tilde\phi$, productivity
      aggregates $Z,\breve Z,\tilde Z$, price indices $P_{djt}$, and revenues
      $R^{trad},R^{new},R^{x}$ satisfy the production, cutoff, price-index, and revenue
      equations of \S\ref{sec:cutoffs} and the trade block, including the $\xi$ factor
      in $\breve Z$ \eqref{eq:Zbreve}.
\item \textbf{Expenditure clearing.} $E_{oj,t}=E^{F}_{oj,t}+E^{M}_{oj,t}+E^{G}_{oj,t}$
      with intermediate demand built from $TFC$.
\item \textbf{Factor markets.} $w_{o,t}$ and $r_{o,t}$ satisfy the labor- and
      land-payment equations of \S\ref{sec:factor} (factor income $=$ true factor cost
      $TFC$ plus, for labor, the fixed-cost bill $\Phi^{F}_o$).
\item \textbf{Government and rebate.} $\bar\tau_t$ and $\bar\pi_t$ satisfy the budget
      formulas of \S\ref{sec:gov} (subsidy outlay $\sum s\,TFC$, infrastructure buying goods).
\item \textbf{Trade balance.} $\mathrm{EX}_t-\mathrm{IM}_t=\mathrm{NX}^{\star}_t$
      \eqref{eq:tradebalance}; equivalently the aggregate resource constraint
      \eqref{eq:walras} holds.
\end{enumerate}

The code treats the static equilibrium as a fixed point in
\[
  x_t = (w_t, r_t, P_t, E_t, \bar\tau_t, \bar\pi_t),
\]
with $L_t$ implied from $x_t$ and $L_{t-1}$.

\section{Dynamic path}
The full dynamic equilibrium is computed sequentially because the only endogenous
state variable is lagged population:
\begin{equation}
  L_{t-1} \longrightarrow \text{solve static equilibrium at } t \longrightarrow L_t.
\end{equation}
For $t=1$, the state is the initial population vector $L_0$. For $t \ge 2$, the
state is the solved population vector from period $t-1$.

\section{Numerical solution}
The solver uses a damped fixed-point iteration:
\begin{enumerate}[leftmargin=1.5em]
\item start from a positive guess for $(w,r,P,E,\bar\tau,\bar\pi)$,
\item compute the implied mapping $F(x_t)$ from the equilibrium equations,
\item update each component with geometric damping in logs,
\item stop when the maximum absolute log residual falls below tolerance.
\end{enumerate}

For example, wages are updated by
\begin{equation}
  w^{new} = \exp\left[(1-\lambda)\log w^{old} + \lambda \log w^{imp}\right],
\end{equation}
and analogously for $r$, $P$, and $E$. Later periods are warm-started from the previous
period's solution.

\paragraph{Sign-unrestricted rebate (review \S2.4).} The rebate $\bar\pi_t$ and (in
principle) net transfers can be \emph{negative}, so they cannot be updated by geometric
damping in logs, which floors them at a small positive number and can report a floored,
non-converged point as a solution. They should be updated in levels --- e.g.\ damping
$1+\bar\pi_t$ geometrically, or $\bar\pi_t$ arithmetically --- and residuals reported at
the true fixed point. The bounded-support $\max\{1,\cdot\}$ truncations in the cutoffs
introduce kinks where a cutoff hits $1$ or $\kappa$; the solver should be checked for
convergence to the genuine fixed point across these kinks rather than to a projected one.
% [REVIEW FLAG \S2.4] This describes the required solver change; the edit to
% solver.py (_geom_update_scalar on taubar/pibar) and equilibrium.py floors is a code-pass
% item.

\section{Restrictions, existence, and market structure}\label{sec:restrictions}
\subsection{Parameter and equilibrium restrictions}
\begin{itemize}[leftmargin=1.5em]
\item $L_0$ is normalized to sum to one (so populations are shares; see the numeraire
      discussion in \S3.1).
\item $\gamma_{oj,t} + \sum_k \gamma^{IO}_{ojk,t} = 1$ and
      $\breve\gamma_{oj,t} + \sum_k \breve\gamma^{IO}_{ojk,t} = 1$ for every $(o,j,t)$.
\item $\breve\gamma_{oj,t}\breve\beta_{oj,t} = \gamma_{oj,t}\beta_{oj,t}$, so the
      land share inside value added is preserved across agricultural technologies.
\item Outside the heavy-manufacturing input, the mechanized agricultural input shares
      equal the baseline input shares.
\item Subsidy rates satisfy $0 \le s_{oj,t} < 1$ (strict, since $1/(1-s)$ appears in
      \eqref{eq:TFC}).
\item \textbf{PIGL regularity (new, review \S7.3):} $\eta>0$, $v_{\mathrm{Agri}}>0$, and the
      income range keeps every share $\psi_{dj,t}\in[0,1]$ at the solution.
\item \textbf{Cutoff ordering (new, review \S2.2):} in agriculture,
      $\bar\phi_{oj,t}\le\breve\phi_{oj,t}\le\tilde\phi_{oj,t}$, so that exporters are
      adopters and no firm is simultaneously a mechanized exporter and a traditional
      domestic producer. This must be verified at the solution, not assumed.
\end{itemize}

\subsection{Existence, uniqueness, and multiplicity (review \S3.10)}\label{sec:existence}
Lagging agglomeration and congestion in $L_{t-1}$ removes within-period increasing returns,
which helps, but the static system still couples IO linkages, selection cutoffs with kinks
(the $\max\{1,\cdot\}$ truncations), and non-homothetic demand; existence and uniqueness
are not automatic and should be addressed (cf.\ Allen--Arkolakis--Li sufficient
conditions). More importantly, the \emph{dynamic} system with $\rho_j>0$ can exhibit path
multiplicity: temporary policy can move the economy across basins of attraction. This is
not a nuisance but the economics of big-push industrial policy (the Choi--Levchenko
question), and a full treatment should characterize when temporary HCI shifts are
permanent.
% [REVIEW FLAG \S3.10] Existence/uniqueness results and the multiplicity analysis are
% research deliverables (propositions), not edits. Recorded here so the primer stops being
% silent about them.

\subsection{Market structure caveat (review \S3.5)}
Agriculture is currently Melitz monopolistic competition: millions of rice farmers each
earning a $\sigma/(\sigma-1)$ markup on a differentiated variety. This is hard to defend
and distorts subsidy pass-through and the adoption calculus. The literature standard is
competitive, homogeneous-good agriculture with heterogeneous \emph{land quality} (Sotelo
2020), so that selection into mechanization is driven by land heterogeneity rather than
monopoly power. At minimum the market-structure assumption must be justified or shown to be
innocuous by robustness.
% [REVIEW FLAG \S3.5] Re-specifying agriculture as competitive is a structural change, not a
% theory edit; flagged for the later phase.

\section{Discrepancies between this note and the current code}\label{sec:discrepancies}
This revision corrects the \emph{theory}. Until the accompanying code pass lands, several
equations here intentionally diverge from \texttt{src/korea\_growth/} (section numbers below
refer to \texttt{docs/model\_review.md}). The outstanding code changes are:
\begin{enumerate}[leftmargin=1.5em]
\item \textbf{$\xi$ in the domestic mechanized aggregate} \eqref{eq:Zbreve}:
      \texttt{trade.py} (the \texttt{Zbreve} line) omits $\xi$; multiply it in. (review \S2.1)
\item \textbf{Subsidy grossing-up}: replace the private cost $\frac{\sigma-1}{\sigma}R$ by
      $TFC=\frac{\sigma-1}{\sigma(1-s)}R$ in factor demands and intermediate demand, and
      use $\sum s\,TFC$ for the subsidy outlay in \texttt{equilibrium.py}. (review \S1.3)
\item \textbf{Land rents to households}: add $r_{d}H_{d}/L_{d}$ to disposable income in
      \texttt{preferences.py}/\texttt{equilibrium.py}. (review \S1.1)
\item \textbf{Fixed costs as labor income}: add $\Phi^{F}_{o}$ to labor demand. (review \S1.2)
\item \textbf{Infrastructure buys goods}: add $E^{G}$ and route $G^{infra}$ into
      goods-market clearing. (review \S1.4)
\item \textbf{Migration real-income index}: use the positive $\mathcal{Y}_{d}$ of
      \eqref{eq:realincome} in place of the sign-ambiguous $V_d$ in
      \texttt{preferences.py}. (review \S2.3)
\item \textbf{Sign-unrestricted $\bar\pi$}: update in levels, not logs. (review \S2.4)
\item \textbf{New checks}: cutoff ordering, PIGL share bounds, and a Walras's-law assertion
      at the solved equilibrium (\texttt{checks.py} / tests). (review \S1, \S2.2, \S7.3)
\end{enumerate}
Items 1--8 are the ``theory-consistent'' minimum. Structural upgrades flagged above ---
capital and credit wedges (review \S3.1), annual/forward-looking migration (review \S3.2),
competitive agriculture (review \S3.5), sector-specific $\sigma_j$ and non-traded services
(review \S3.6), separate inbound/outbound port frictions (review \S3.7), county-level
geography and estimation (review \S4) --- are deliberately \emph{not} attempted in this pass.

\section{Interpretation}
The model combines five ingredients:
\begin{enumerate}[leftmargin=1.5em]
\item non-homothetic demand and endogenous migration,
\item Melitz-style firm heterogeneity with bounded Pareto draws,
\item input-output linkages across sectors,
\item sector-specific agglomeration from lagged population,
\item policy wedges acting through productivity, fixed costs, subsidies, trade costs,
      migration costs, and foreign demand.
\end{enumerate}

That is the economic system solved by the repository. Any empirical exercise or policy
counterfactual run through the package is a perturbation of this equilibrium system.

\end{document}
